### 2014-12-24 qfenv_paper/ApJ submission/cover_apj.tex
draft for submission

\documentclass{letter}
\usepackage{setspace}

\begin{document}

\begin{flushleft}
\begin{singlespace}
Department of Physics\\
New York University\\
4 Washington Place, Room 424\\
New York, NY 10003
\end{singlespace}

\vspace{0.5cm}

Dec 22, 2014

\vspace{0.5cm}

Dear Editor,

Please find enclosed our paper "PRIMUS: Galaxy Environment on the Quiescent Fraction Evolution at \lowercase{\textit{z}}$<0.8$," which we are submitting exclusively to the Astrophysical Journal.  Please contact me for any correspondence concerning our manuscript, my email address is \texttt{chh327@nyu.edu}.

\vspace{0.5cm}

Sincerely,

ChangHoon Hahn
\end{flushleft}

\end{document}