### 2015-03-20 qfenv_paper/ApJ submission/referee_response.txt
referee responses almost done

Dear Editor,

We thank the referee for the thorough review of our work and for the constructive report that improved the quality of the paper.

Please find below responses to the referee’s comments for the submission along with a number of changes made to our paper.

Thank you, 
ChangHoon Hahn

-
ChangHoon Hahn
Center for Cosmology and Particle Physics
New York University 
4 Washington Place 
New York, NY 10003
chh327@nyu.edu

================================================================================
        February 2, 2015

        Mr. ChangHoon Hahn
        Center for Cosmology and Particle Physics
        New York University
        4 Washington Place
        New York, NY 10003

        Title: PRIMUS: Galaxy Environment on the Quiescent Fraction Evolution at z < 0.8

        Dear Mr. Hahn,

        I have received the referee's report on your above submission to The
        Astrophysical Journal, and appended it below. As you will see, the
        referee thinks that your article is interesting and that it will merit
        publication once you have addressed the issues raised in the
        report.

        When you resubmit the manuscript, please include a detailed cover
        letter containing the (mandatory) listing of the changes you've made to
        the text and your responses to the report.

        Click the link below to upload your revised manuscript;         
        http://apj.msubmit.net/cgi-bin/main.plex?el=A6Ew1TL6A3CfGS1J7A9ftdhSCRuGAD5aB2tFkIbuQxBwZ
        Alternatively, you may also log into your account at the EJ Press web
        site, http://apj.msubmit.net. Please use your user's login name:
        chahah. You can then ask for a new password via the Unknown/Forgotten
        Password link if you have forgotten your password.

        Reviewers find it helpful if the changes in the text of the manuscript
        are easily distinguishable from the rest of the text. Therefore we ask
        you to print changes in bold face. The highlighting can be removed
        easily after the review.

        The Astrophysical Journal has adopted a new policy that manuscript
        files become inactive, and are considered to have been withdrawn, six
        months after the most recent referee's report goes to the authors,
        provided a revised version has not been received by that time.

        If you have any questions, feel free to contact me.

        Sincerely,
        Lee Armus
        ApJ Scientific Editor
        California Institute of Technology
        apjla@ipac.caltech.edu
        ---------------------------------------------------------------------- 
        Referee Report 
        Reviewer's Comments:
        Report on 'PRIMUS: Galaxy Environment on the Quiescent Fraction
        Evolution at z < 0.8' by Hahn et al.

        The authors have used the large PRIMUS redshift survey to
        determine the quiescent fraction as a function of environment,
        mass and redshift, with SDSS used for the lowest redshift
        bin. This is topical as it pertains to discussions of how
        galaxies become quiescent - often called quenching. This study is
        new in that it utilizes the largest redshift survey over the
        range 0.2 < z < 0.8. Having said that, the drawback of this
        particular survey is that the cosmological redshift uncertainties
        are significantly larger than one would expect from peculiar
        velocities because of the low-dispersion spectra used. This
        increases the noise and systematic uncertainties in quantifying
        environments. Overall the paper is sufficiently clear and
        original to warrant publication. I would recommend some further
        investigation into the environment measures and whether this
        might introduce some systematic trends, as well as some minor
        changes.

        Major comments:

        The environmental measures make use of long cylinders, with the
        length along the redshift direction. The height is noted as being
        35 Mpc. This may actually be +-35 Mpc. At z=0.7, an average
        redshift uncertainty of 0.005*(1+0.7) corresponds to 25 Mpc. The
        half height of the cylinder then corresponds to 1.4 sigma (or
        only 0.7 sigma if the cylinder is +- 17.5 Mpc). The authors
        should investigate whether the uncertainties of the PRIMUS
        redshifts vary as a function of redshift, and whether the
        uncertainties depend on galaxy type (star forming or
        quiescent). This potentially affects the quiescent fraction as a
        function of environment as certain types of galaxies might be
        more likely to be scattered out of higher density environments.

RESPONSE: 
The height represents the full height of the cylindrical aperture (+/— 17.5
Mpc). We select the dimensions to minimize aperture height while accounting for
the redshift uncertainties, especially at our sample's highest redshift bin z~0.7.

Cool et al. 2013 specifically compare redshift accuracy and success rate of
PRIMUS galaxies to high resolution spectroscopic redshifts from COSMOS, DEEP2,
and VVDS. They do not note any significant redshift dependence or dependence
with observed-frame color. We now explicitly refer to Cool et al. (2010) in
Section 2.1.

In addition, Moustakas et al. 2013 emphasize in their Section 3.1 after
extensive comparison to high resolution spectroscopic redshift that the
redshift success in PRIMUS is not a strong function of galaxy type. Also, when
we compare a subsample of the data used in our analysis to high resolution
spectroscopic redshifts we find no significant dependence on galaxy type or
redshift (no significant variations in the dispersion of delta z/(1+z) or the
outlier rate between quiescent and star-forming populations).

        Even though the lowest masses are generally not discussed in the
        paper, there is a concern that the galaxies at 10^9.5 and 10^10
        solar masses show little dependence on environment in the PRIMUS
        sample, especially over redshifts 0.2-0.4. Other authors have
        shown that the environmental dependence is getting larger below
        10^10.5. The SDSS data cutoff at 10^10.4 in Figure 4. The authors
        could look at a 0.02 to 0.05 redshift sample to test whether the
        SDSS trend continues and shows quite a different trend to the
        PRIMUS 0.2-0.4 sample as suggested by the high-density
        environment data presented in Figure 4. One would not expect a
        big change in behaviour between z=0 and z=0.3, and one
        explanation for the difference could be the redshift
        uncertainties in PRIMUS and how they depend on galaxy
        type. Noting that the SDSS uncertainties that have been
        artificially included are 0.005 in 1+z with a Gaussian
        distribution, which is unlikely to have exactly the same form as
        the PRIMUS redshift uncertainties.

RESPONSE: 
We added a discussion on the lack of environment dependence in the quiescent
fraction at z ~ 0.3 for M* < 10^10 solar masses near the end of Section 4.1.

Although we observe little environment dependence at these redshifts and mass
range, we note that the uncertainties calculated from our jack-knife technique
underestimate the total uncertainty contribution from cosmic variance.

At these M* < 10^10 solar masses, only three out of the five targeted PRIMUS
Fields (XMM-SXDS, XMM-CFHTLS, and COSMOS) contribute galaxies, with XMM-SXDS
and XMM-CFHTLS being spatially adjacent. A rough estimate of the cosmic
variances for these three fields using the expressions derived in Driver et al.
(2010) put cosmic variance uncertainties > 60%. This estimate does not take
into account the unique geometry of the PRIMUS fields or the edge-cuts we
impose, which only increase the cosmic variance uncertainties. Therefore, the
quiescent fraction comparison at z ~ 0.3 for M* < 10^10 solar masses cannot be
accurately made using our PRIMUS sample.

A comparison to the low mass SDSS quiescent fraction is currently made
difficult by the mass-to-light ratio and the bright and faint limits of the
SDSS data. Future investigation, which will correct for the photometric biases
of the SDSS data will also extend the low mass limit of the quiescent fraction
analysis using data from the NASA Sloan Atlas Catalog.

        Note on above, I think it is fine for this paper to focus on
        fiducial points of 10^10.5 and 10^11 solar masses, it is just
        that looking to lower masses may indicate a problem with the
        methodology.

        Minor comments:

        Last sentence of abstract: 'However' is the wrong joining term; 'In
        other words' would be more appropriate.

‘However’ has been corrected to ‘In other words’.

        Section 2.1: How does iSEDFIT deal with dust obscuration? How
        successful might it be? A couple or more sentences on this would be
        useful.

We now explicitly mention that the time-dependent dust attention curve of
Charlot & Fall (2000) is used by iSEDfit to deal with dust obscurations in
Section 2.1. We also now explicitly refer reads to Moustakas et al. (2013) for
further details on the effects of prior parameter choices on the extracted
physical properties of galaxies.

        Section 2.3: GALEX does not provide redshifts. When referring to
        redshifts, they should not be called SDSS-GALEX redshifts. It is only
        an SDSS-GALEX sample.

We now correctly specify that the redshifts from the SDSS-GALEX sample are SDSS
redshifts in Section 2.4.

        Section 2.5: State whether 35 Mpc is half height or full height of
        cylinder. I presume these are comoving Mpc, state this.

We now specifically state that H_ap is the full height of the cylinder and that
the aperture dimensions are comoving distances in Section 2.5.

        There is perhaps surprisingly no evolution in M_{r,lim}. Has there been
        an evolution correction applied for M_r?

No, we do not apply any evolution correction to the absolute magnitude values.

        N_{env}. Can this have a value between 0 and 1? I would have thought
        all the weights would be >=1. Therefore better to state the low-density
        environment as n_{env} = 0.

No, n_{env} cannot have a value between 0 and 1. We have corrected the
low-density environment cut-off to n_{env} = 0 throughout the text and in all
the figures.

        Section 2.6: Why does this need to use random points? Why not a finely
        spaced regular grid in RA and DEC?

We use random points due to the convenience of the ransack commands, which
generates random points given a survey geometry. We believe that the referee's
suggestion of a finely spaced regulard grid would not make a significant
difference in accounting for the edge effects.

        End of page 9: missing text after $\sim0.3$ to $\sim0.7$ [for high density].

Corrected.

        Page 12 and elsewhere: The caveat that PRIMUS has more galaxies, and
        therefore better statistics than zCOSMOS is that the environmental
        measures are possibly significantly less reliable because of redshift
        uncertainties. I think this should be pointed out a couple of more
        times than is done currently within the text. There is also the issue
        of overall redshift completeness (targeting and redshift success
        combined), how does PRIMUS compare with zCOSMOS in this regard?

We now explicitly state, in Section 4.3, that our galaxy environment measures
are susceptible to contamination due to PRIMUS redshift uncertainties. We also
emphasize the role of the redshift uncertainties in motivating the more
stringent high density environment classifications in Section 4.2.

As suggested by referee, we also discuss the overall redshift completeness
of PRIMUS versus zCOSMOS in Section 4.3.

        Section 5, item 3: Note the mass range over which the results are valid.

We now explicitly state that the mass range (10^10.5-10^11.0 M_sun) where the
results are valid.

------------------------------------------------------------------------------------
OTHER CHANGES:

- Title has been changed to “Effects of Galaxy Environment on the Quiescent
  Fraction Evolution at z < 0.8”

- Typo in Section 2.1 of one of the PRIMUS fields has been corrected.