### 2015-04-20 qfenv_paper/ApJ submission/referee_response2.txt
Referee Report #2

Dear Editor,

We once again thank the referee for the thorough review of our work and
for the constructive report.

Please find below responses to the referee's comment along with changes
made to our paper.

Thank you, 
ChangHoon Hahn

================================================================================
        April 11, 2015

        Mr. ChangHoon Hahn
        Center for Cosmology and Particle Physics
        New York University
        4 Washington Place
        New York, NY 10003

        Title: PRIMUS: Effects of Galaxy Environment on the Quiescent
        Fraction Evolution at z < 0.8, ApJ97613R1

        Dear Mr. Hahn,

        The revised version of your paper cited above was sent back to
        the referee, who is mostly satisfied with the changes you have
        made. He/she just has a few remaining comments (appended below),
        which I would like you to take into consideration.

        When you resubmit the manuscript, please include a detailed
        cover letter containing the (mandatory) listing of the changes
        you've made to the text and your responses to the report.

        Click the link below to upload your revised manuscript; 
        http://apj.msubmit.net/cgi-bin/main.plex?el=A1Ew7DXK1A4CfGS3J7A9ftduFQ0I1nKsyQ3P8J1nY5FIQZ
        Alternatively, you may also log into your account at the EJ
        Press web site, http://apj.msubmit.net. Please use your user's
        login name: chahah. You can then ask for a new password via the
        Unknown/Forgotten Password link if you have forgotten your
        password.

        Reviewers find it helpful if the changes in the text of the
        manuscript are easily distinguishable from the rest of the text.
        Therefore we ask you to print changes in bold face. The
        highlighting can be removed easily after the review.

        The Astrophysical Journal has adopted a new policy that
        manuscript files become inactive, and are considered to have
        been withdrawn, six months after the most recent referee's
        report goes to the authors, provided a revised version has not
        been received by that time.

        If you have any questions, feel free to contact me.

        Sincerely,
        Lee Armus
        ApJ Scientific Editor
        California Institute of Technology
        apjla@ipac.caltech.edu

        ---------------------------------------------------------------------- 
        Referee Report
        Reviewer's Comments:

        I have looked at the changes. I think the argument about the
        cosmic variance affecting the quiescent fraction is overstated -
        some of it irrelevant. The cosmic variance has a much larger
        effect on the number counts than on the fraction of quiescent
        galaxies in high density environments; the latter is not
        estimated by cosmic variance calculators/equations. I agree the
        jack-knife method may underestimate the errors. Alternatively
        perhaps the stellar mass limits are not strict enough when
        estimating the quiescent fraction, since the 'missed galaxies'
        will preferentially be quiescent. I recommend the authors think
        about this a bit more; or at least remove the text about the
        cosmic variance. This should equally affect the >10^10 sample
        for the z=0.3 bin.

RESPONSE: 
We have removed the text discussing the cosmic variance.

        You now undersell PRIMUS versus zCOSMOS. The targeting
        completeness for zCOSMOS is much lower, see Fig. 5 of Lilly et
        al. 2009 for the 10k sample. What is most relevant is combined
        targeting completeness and redshift success - even a rough
        estimate for zCOSMOS would be a useful comparison - and would be
        a favourable metric for PRIMUS I think.

RESPONSE: 
We now explicitly mention the targeting completeness (spatial sampling
rate) of zCOSMOS in addition to the redshift completeness and list an
estimate of the overall completeness rate of zCOSMOS mentioned in Knobel
et al. (2012).

--------------------------------------------------------------------------------
OTHER CHANGES:

- We now state that Iovino et al. (2010) remark that the uncertainties
  in the quiescent fraction fits are too large to make any conclusions
  about the quiescent fraction evolution.